Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-06-17 forecast/paper/paper.tex
Typo fixes, update revision fig

Finally, it is worth noting that examining the importance of additional features
for prediction is a core question in inferential statistics and econometrics,
with work dating back to at least \cite{Granger:1969}. Still today, drawing
rigorous inference based on predictions, without (or with lean) assumptions, is
an active field of research from both the applied and theoretical angles;
see, e.g., \cite{Diebold:2002, McCraken:2007, Diebold:2015, Stokes:2017,
  Lei:2018, Rinaldo:2019, Williamsom:2020, Zhang:2020, Dai:2021, Fryer:2021}.
Our take on this problem is in line with much of this literature; however, in
order to avoid making any explicit assumptions, we do attempt to make explicit
significance statements, and instead, broadly examine the stability of our
conclusions with respect to numerous modes of analysis.

@@

In this paper, all models are fit with ``vintage'' training data. This means
that for a given prediction date, say, September 1, 2020, we train models 
using only the data that would have been available to us \textit{as of}
September 1 (imagine that we ``rewind'' the clock to September 1, and query the
COVIDcast API to get the latest data it would have had available at that point
in time.)  This is made possible by the COVIDcast API's comprehensive data  
versioning system (described in more detail in our companion paper).  We also
use the \texttt{evalcast} R package \cite{EvalcastR}, which streamlines the
process of training arbitrary prediction models over a sequence of prediction
dates, by constructing the proper sequence of vintage training data sets.

\begin{figure}[tb!]
  \includegraphics[width=\columnwidth]{fig/revisions-dv-jhu-1.pdf}
  \caption{Revision behavior for signals in the New York City HRR from two
    sources. Each colored line corresponds to the data as reported on that
    date. DV-CLI is revised upward throughout the month of December, though the
    amount fades as we look further back in time. In contrast, JHU-CSSE data
    remains ``as reported'' on December 3 until a major correction is made on
    December 31, affecting all data over the period. }
  \label{fig:vintage}
\end{figure}

@@

In addition to the above ``pure'' autoregressive model, we consider five
probabilistic forecasting models of the form: 
\begin{multline}
\label{eq:forecast_ar_x}
\Quantile_\tau(Y_{\ell,t+a} \,|\, Y_{\ell,s}, X_{\ell,s}, s \leq t)  
= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j} + 
\sum_{j=0}^2 \gamma^{a,\tau}_j X_{\ell,t-7j},
\end{multline}
where $X_{\ell,t}$ denotes any one of the five auxiliary indicators---CHNG-CLI, 
CHNG-COVID, CTIS-CLI-in-community, DV-CLI, or Google-AA---at location $\ell$ and
time $t$. Note that we apply the same lags (current value, along with the values
7 and 14 days ago) for the auxiliary indicators as we do for the case
rates. Training then proceeds just as before: we use the same 7 quantile levels
in \eqref{eq:quantile_levels}, and fit quantile regression separately for each 
level $\tau$, using data from all 306 HRRs and a trailing window of 21 days of
training data.

@@

In addition to this ``pure'' autoregressive model, we also consider five
logistic models of the form:  
\begin{multline}
\label{eq:hotspot_ar_x}
\logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s}, X_{\ell,s}, s \leq t) \big)
= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j} +  
\sum_{j=0}^2 \gamma^{a,\tau}_j X^\Delta_{\ell,t-7j},
\end{multline}
where we use \smash{$X^\Delta_{\ell,t} = (X_{\ell,t} - X_{\ell,
    t-7})/(X_{\ell,t-7})$}, and again $X_{\ell,t}$ stands for any of the five  
auxiliary indicators at location $\ell$ and time $t$.  We fit all models
\eqref{eq:hotspot_ar}, \eqref{eq:hotspot_ar_x} using logistic regression, 
pooling all 306 HRRs and using a 21-day trailing window for the training data.

@@

Figure~\ref{fig:forecast} displays the results for forecasting, stratified by
ahead value and averaged over all HRRs and forecast dates (June 9--December 31,
2020).  Shown is the average WIS for each forecast model divided by the averaged
WIS from a ``baseline'' model, which is basically a flat-line forecaster (its
median forecast for $Y_{\ell,t+a}$ is always $Y_{\ell,t}$, with predicted
quantiles defined around this flat line based on historical variation). This is
the same baseline model as that used in the COVID-19 Forecast Hub.  Here, we  
divide by the average WIS of the baseline model in order to put the y-axis on an 
interpretable scale.  We can see that all curves are well below 1, which means
that all of the models, including AR, outperform the baseline on average over
the forecast period (though their advantage over the baseline clearly narrows as
the ahead value grows).

@@

Figure~\ref{fig:hotspot} displays the results for hotspot prediction, again
stratified by ahead value and averaged over all HRRs and prediction dates (June 
16--December 31, 2020).  We can see many similarities to the forecasting
results: CHNG-COVID and DV-CLI offer the biggest improvement over AR, all models
including AR degrade in performance towards the baseline (this is now just a
classifier based on random guessing, which achieves an AUC of 0.5) as the ahead
values grow.  However, one clear difference is that the CHNG-CLI model performs
quite well in hotspot prediction, close to the best-performing indicator models
for many of the ahead values. [TODO: should we insert one line explaining
this---you're somewhat protected form volatiliy in a logistic model thanks to
the sigmoid function used to define the probabilities?]

@@

One might ask if the benefits we observe in forecasting and hotspot
prediction have
anything to do with the actual indicator themselves; could it instead be that
the indicators are simply providing ``implicit regularization'' on top of the
basic AR model, when we include them as lagged features in
\eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar_x}?

>>>>

Finally, it is worth noting that examining the importance of additional features
for prediction is a core question in inferential statistics and econometrics,
with work dating back to at least \cite{Granger:1969}. Still today, drawing
rigorous inference based on predictions, without (or with lean) assumptions, is
an active field of research from both the applied and theoretical angles;
see, e.g., \cite{Diebold:2002, McCraken:2007, Diebold:2015, Stokes:2017,
  Lei:2018, Rinaldo:2019, Williamsom:2020, Zhang:2020, Dai:2021, Fryer:2021}.
Our take on this problem is in line with much of this literature; however, in
order to avoid making any explicit assumptions, we do not attempt to make formal 
significance statements, and instead, broadly examine the stability of our
conclusions with respect to numerous modes of analysis.

@@

In this paper, all models are fit with ``vintage'' training data. This means
that for a given prediction date, say, September 1, 2020, we train models 
using only the data that would have been available to us \textit{as of}
September 1 (imagine that we can ``rewind'' the clock to September 1, and query
the COVIDcast API to get the latest data it would have had available at that
point in time.)  This is made possible by the COVIDcast API's comprehensive data   
versioning system (described in more detail in our companion paper).  We also
use the \texttt{evalcast} R package \cite{EvalcastR}, which streamlines the
process of training arbitrary prediction models over a sequence of prediction
dates, by constructing the proper sequence of vintage training data sets.

\begin{figure}[tb!]
  \includegraphics[width=\columnwidth]{fig/revisions.pdf}
  \caption{Revision behavior for two signals in the HRR containing New York City
    HRR.  Each colored line corresponds to the data as reported on a particular
    date (\textit{as of} dates varying from September 28 through October
    19). The left panel shows the DV-CLI signal, which was regularly revised
    throughout the period, though the effects fade as we look further back in   
    time. In contrast, the right panel shows case rates reported by JHU CSSE,
    which remain ``as reported'' on September 28, with a huge spike towards the 
    end of this period, until a major correction is made on October 19, which
    brings down this spike and affects all prior data as well.}
  \label{fig:vintage}
\end{figure}

@@

In addition to this pure autoregressive model, we also consider five
probabilistic forecasting models of the form:  
\begin{multline}
\label{eq:forecast_ar_x}
\Quantile_\tau(Y_{\ell,t+a} \,|\, Y_{\ell,s}, X_{\ell,s}, s \leq t)  
= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j} + 
\sum_{j=0}^2 \gamma^{a,\tau}_j X_{\ell,t-7j},
\end{multline}
where $X_{\ell,t}$ denotes any one of the five auxiliary indicators---CHNG-CLI, 
CHNG-COVID, CTIS-CLI-in-community, DV-CLI, or Google-AA---at location $\ell$ and
time $t$. Note that we apply the same lags (current value, along with the values
7 and 14 days ago) for the auxiliary indicators as we do for the case
rates. Training then proceeds just as before: we use the same 7 quantile levels
in \eqref{eq:quantile_levels}, and fit quantile regression separately for each 
level $\tau$, using data from all 306 HRRs and a trailing window of 21 days of
training data.

@@

In addition to this pure autoregressive model, we also consider five
logistic models of the form:  
\begin{multline}
\label{eq:hotspot_ar_x}
\logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s}, X_{\ell,s}, s \leq t) \big)
= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j} +  
\sum_{j=0}^2 \gamma^{a,\tau}_j X^\Delta_{\ell,t-7j},
\end{multline}
where we use \smash{$X^\Delta_{\ell,t} = (X_{\ell,t} - X_{\ell,
    t-7})/(X_{\ell,t-7})$}, and again $X_{\ell,t}$ stands for any of the five  
auxiliary indicators at location $\ell$ and time $t$.  We fit all models
\eqref{eq:hotspot_ar}, \eqref{eq:hotspot_ar_x} using logistic regression, 
pooling all 306 HRRs and using a 21-day trailing window for the training data.

@@

Figure~\ref{fig:forecast} displays the results for forecasting, stratified by
ahead value and averaged over all HRRs and forecast dates (June 9--December 31,
2020).  Shown is the average WIS for each forecast model divided by the averaged
WIS from a baseline model, which is basically a flat-line forecaster (its
median forecast for $Y_{\ell,t+a}$ is always $Y_{\ell,t}$, with predicted
quantiles defined around this flat line based on historical variation). This is
the same baseline model as that used in the COVID-19 Forecast Hub.  Here, we  
divide by the average WIS of the baseline model in order to put the y-axis on an 
interpretable scale.  We can see that all curves are well below 1, which means
(smaller WIS is better) that all of the models, including AR, outperform the
baseline on average over the forecast period (though this advantage narrows as
the ahead value grows).

@@

Figure~\ref{fig:hotspot} displays the results for hotspot prediction, again
stratified by ahead value and averaged over all HRRs and prediction dates (June 
16--December 31, 2020).  We can see many similarities to the forecasting
results (recall larger AUC is better): CHNG-COVID and DV-CLI offer the biggest
improvement over AR, all models including AR degrade in performance towards the
baseline (this is now just a classifier based on random guessing, which achieves
an AUC of 0.5) as the ahead values grow.  However, one clear difference is that
the CHNG-CLI model performs quite well in hotspot prediction, close to the
best-performing indicator models for many of the ahead values.

[TODO: should we
insert one line explaining this---you're somewhat protected form volatiliy in a
logistic model thanks to 
the sigmoid function used to define the probabilities?]

@@

One might ask if the benefits we observe in forecasting and hotspot
prediction have
anything to do with the actual indicator themselves; could it instead be that
the indicators are simply providing \textit{implicit regularization} on top of 
the basic AR model, when we include them as lagged features in
\eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar_x}?
